Lattice Boltzmann simulations of magnetically responsive\\ anisotropic particle stabilized emulsions

### 2025-02-16 chapters/paper1.tex
Changes 1. Fixed reference formatting in bib file and added stylesheet for rendering 2. Modified formatting to improve human readability in an IDE
Changes 1. Made tex files human readable again 2. Added content to paper3 tex file 3. Added licenses to some of the images from other papers
Changes 1. Added more analysis to application of magnetic fields during formation chapter 2. Added references to fix some (not all) literature reference issues 3. Added some detail in analysis techniques section

Enter some blurb here for a short introduction to content

\subsubsection{Model parameters and simulation setup}\label{sec:parameters}

The dynamics of the binary fluid mixture is governed by inertial,
viscous, and surface tension forces. The relative magnitudes of these
forces at a characteristic length scale \(L\) and characteristic
velocity \(U\) are captured by the capillary number $\mathrm{Ca}=\frac{\eta U}{\sigma}$ and the Weber number
\(\mathrm{We}=\frac{\rho U^2 L}{\sigma}\). Here we are primarily interested in spinodal decomposition of the mixture, where we analyze the dynamics after the initial formation of the interfaces and neglect residual diffusion. Hence the Peclet number \(\mathrm{Pe}=\frac{UL}{\Gamma}\) is of lesser importance and we do not consider it when choosing our simulation parameters.

The equations of motion of our model comprise the relevant materials
properties: the density \(\rho\), viscosity \(\nu\), and surface tension \(\sigma\) of the binary fluid, and the mass \(m_p\), size \(R_p\) and aspect ratio \(\alpha\), and dipole moment \(m\) of the particles. In addition, we can control the particle volume fraction \(\phi_p\) and
magnetic field strength \(B\). These nine quantities involve four unit
dimensions (mass, length, time, and charge), hence we can use five
dimensionless variables to specify the parameters of the system. These
are the particle aspect ratio $\alpha$, particle volume fraction
\(\phi_p\), particle vs.~fluid density ratio
$\xi = \frac{3m_p}{4\pi \rho V_p}$, nominal Weber number $\mathrm{We} = \frac{\rho \sigma V_p^{1/3}}{\eta^2}$, and magnetic Bond
number \(\bar{B} = m B/(\sigma A_p)\), where
\(V_p=(4/3)\pi R_\parallel^3/\alpha^2\) is the particle volume and
\(A_p=\max\left( \pi R_\parallel^2/\alpha, \pi R_\parallel^2/\alpha^2 \right)\)
the area of the larger cross-section of the ellpsoidal particles.

For our simulations, we choose units \(\Delta x\), \(\Delta t\),
\(\Delta m\), and \(\Delta i\) such that \(V_p=\hat{V}_p(\Delta x)^3\),
\(\nu=\hat{\nu}(\Delta x)^2/\Delta t\),
\(\sigma=\hat{\sigma}\Delta m/(\Delta t)^2\), and
\(m=\hat{m}\Delta i(\Delta x)^2\). We used \(V_p=2000\pi/3\),
\(\hat{\nu}=1/6\), \(\hat{\sigma}=0.0267\), \(\hat{m}=1\), and a density
ratio \(\xi=1/\hat{\rho}\). For reference, the values of the model
parameters are listed in table \ref{tab:parameters}. The values of
\(\alpha\) and \(\bar{B}\) are variable and are reported
with the results in section \ref{sec:results}.

%\new{We note that the dynamics of the simulations are determined by
%the dimensionless numbers, and the actual unit values $\Delta x$,
%$\Delta t$, $\Delta m$, and $\Delta j$ are arbitrary and only become
%meaningful when a connection to an experimental system is made (see
%below).}

\begin{table}
\centering
\caption{Summary of the parameters of the numerical model and the values used in the simulations}
\label{tab:parameters}
\begin{tabular}{|c|r|}
\hline
Parameter & Value \\
\hline
$L_V/\Delta x$ & 256 \\
$\rho \cdot (\Delta x)^3/\Delta m$ & 0.7 \\
$\nu \cdot \Delta t/(\Delta x)^2$ & 1/6 \\
$\sigma \cdot (\Delta t)^2/\Delta m$ & 0.0267 \\
$\tau/\Delta t$ & 1 \\
$g \cdot \Delta m(\Delta t)^2/(\Delta x)^5$ & 0.08 \\
$\phi_f$ & 0.5 \\
$\phi_p$ & 0.15 \\
$n_p$ & 1200 \\
$\rho_p \cdot (\Delta x)^3/\Delta m$ & 1 \\
$V_p/(\Delta x)^3$ & $2000\pi/3$ \\
$m/(\Delta i(\Delta x)^2)$ & 1 \\
$d_c/\Delta x$ & 2/3 \\
$K_H\cdot (\Delta x)^{1/2}(\Delta t)^2/\Delta m$ & 100 \\
\hline
\end{tabular}
\end{table}

We performed simulations of a binary fluid using the software package
LB3D \cite{schmieschek_lb3d_2017} that implements the lattice Boltzmann method described in section
\ref{sec:methods}. The multicomponent
lattice Boltzmann model is solved in a cubic box of size
\(L_V=256\Delta x\) with periodic boundary conditions. The simulation
box is filled with equal volume fractions \(\phi_f=0.5\) of two fluids,
initially homogeneously mixed with a density
\(\rho=0.7\Delta m/(\Delta x)^3\). The BGK relaxation time is set to
\(\tau=\Delta t\) and the Shan-Chen interaction strength is set to
\(g_{kk^\star}=0.08(\Delta x)^2/(\Delta m(\Delta t)^2)\). We performed
preliminary simulations of a spherical droplet with these parameters and
fitted the pressure difference inside and outside the droplet to the
Young-Laplace law, which yielded a surface tension
\(\sigma=0.0267\Delta m/(\Delta t)^2\). To study the influence of
anisotropic particle shapes on the bijel formation, we performed
simulations for particles with three different aspect ratios:
\(\alpha=1\) for spherical particles, \(\alpha=2\) for prolate
ellipsoids, and \(\alpha=1/2\) for oblate ellipsoids. To match the
particle volume fraction \(\phi_p\) between the different particle
shapes, we kept the volume of the particles fixed and used the same
number of particles for each aspect ratio. The radius along the symmetry
axis of the particles was calculated from
\(V_p=(4/3)\pi R_\parallel^3/\alpha^2\), yielding
\(R_\parallel=7.9\Delta x\) for spheres, \(R_\parallel=12.6\Delta x\)
for prolate ellipsoids, and \(R_\parallel=5\Delta x\) for oblate
ellipsoids, respectively. Particles were added to the fluid by placing
them randomly inside the box. The lubrication cutoff distance was set to
the value \(d_c=(2/3)\Delta x\) recommended by Ladd
\cite{ladd_lattice-boltzmann_2001}, and the strength of the Hertz
potential was set to \(K_H=100\Delta m/((\Delta x)^{1/2}(\Delta t)^2)\).
We first equilibrated the particle positions and orientations by
evolving only the equations of motion of the particles, while the fluid
remained in its initial configuration, until the minimum particle
distance exceeded \(1.2\cdot\max(R_\parallel,R_\parallel/\alpha)\).
After the equilibration, the magnetic flux density
\(\vec{B}=B\hat{\vec{z}}\) was switched on and the full system was
evolved for \(10^5\) timesteps. In the course of the simulation, the binary mixture undergoes spinodal decomposition. The particles adsorb at the coarsening interface and eventually become jammed, resulting in the bicontinuous phase morphology of the bijel.

It is worth discussing briefly the possible realization of our model in
experiments. A common choice for bijel formation in binary liquids is a
mixture of water and lutidine
\cite{clegg_emulsification_2007,herzig_bicontinuous_2007}. The surface
tension between the phases of a water-lutidine system at
\(40^\circ\mathrm{C}\) is around \(\sigma=0.22\ \mathrm{mN/m}\) and the
dynamic viscosity of the lutidine-rich phase is around
\(\eta=2.38\ \mathrm{mPa\,s}\) \cite{grattoni_lower_1993}. Ellipsoidal
particles with various aspect ratios can be formed by mechanically
stretching spherical particles \cite{trevenen_gradient_2021}. Such
particles can be magnetically functionalized by e-beam deposition of a
Nickel layer. Fei et
al.~\cite{fei_magneto-capillary_2018,fei_magneto-capillary_2020}
fabricated coated \(4\ \mathrm{\mu m}\) polystyrene particles with a
permanent magnetic dipole moment of
\(m\approx3\cdot10^{-14}\ \mathrm{Am^2}\). It was shown that the dipole moment is oriented in the direction parallel to the coated interface \cite{yan_linking_2012,yan_rotating_2014}, i.e., it can be aligned with the axis of ellipsoidal particles.
For the given surface tension, the capillary torque on a particle adsorbed at an interface is approximately
$8.8\cdot10^{-16}\ \mathrm{Nm}$. We thus can estimate that a magnetic
flux density of \(30\ \mathrm{mT}\) is able to exert a magnetic torque
that exceeds the capillary torque.
We note that under these conditions,
the dipole-dipole interaction is on the order of
\(10^{-16}\ \mathrm{J}\) and an order of magnitude smaller than the
particle-interface interaction which is of order
\(10^{-15}\ \mathrm{J}\). %Unlike previous works
%\cite{xie_direct_2017,xie_controllable_2021}, we therefore cannot
%neglect the magnetic dipole-dipole interactions between particles.For these bijel properties and the parameters given in
table \ref{tab:parameters}, the spatial resolution of our
simulations is $\Delta x\approx252\ \mathrm{nm}$, and the time step
is $\Delta t=624\ \mathrm{ns}$. This corresponds to a system side length of $L_V \approx 64.5\ \mathrm{\mu m}$ and a runtime of
$T \approx 62.4\ \mathrm{ms}$.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/results/paper1/microstructure_viz.png}
\caption{Snapshots of emulsion gels stabilized by prolate ellipsoids after $t=5\cdot10^3$ timesteps (left) and after $10^5$ timesteps (right). The top row shows the structure forming without a magnetic field. The bottom row shows the structure forming in an applied magnetic field of reduced strength $\bar{B}=1.0$. The snapshots are colored according to the order parameter $\phi$.}
\label{fig:microstructure_viz}
\end{figure}

@@

@@

In LB simulations of spinodal decomposition
\cite{kendon_3d_1999,kendon_inertial_2001}, a characteristic length
scale can be obtained from the structure factor
%
\begin{equation}
S(\vec{k},t) = \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) .
\end{equation}
%
where \(\tilde{\phi}(\vec{k},t)\) is the Fourier
transform of the order parameter fluctuations
\(\phi(\vec{x},t)-\left\langle\phi\right\rangle\). A common measure of
domain size uses the spherically averaged structure factor
%
\begin{equation}
S(k,t) = \frac{1}{n_k} \sum_{k-\frac{\Delta}{2}<k<k+\frac{\Delta}{2}} \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) ,
\end{equation}
%
where \(k=|\vec{k}|\) denotes the modulus of \(\vec{k}\)
and \(n_k\) is the number of lattice sites in a shell of radius \(k\)
and thickness \(\Delta=2\pi/L_V\). The average domain size can then be
defined in terms of the \(n\)-th moment of the structure factor
\cite{laradji_molecular_1996}

\begin{equation}
L_n(t) = 2 \pi (\left( \frac{\sum_k S(k, t)}{\sum_k k^n S(k, t)} )\right)^{\frac{1}{n}}
\end{equation}

It is worth noting that this definition of average domain
size does not necessarily yield values that coincide with the typical
domain size observed visually in snapshots such as those shown in
Fig.~\ref{fig:microstructure_viz}. However, in a dynamic
scaling regime, different measures of domain size are expected to differ
only by a constant ratio. We found that the commonly used measure
\(L_1(t)\) tends to overestimate the observed domain size. Numerically,
the calculation of the spherically averaged structure factor is subject
to poor statistics in the low \(k\)-shells, where the average
\(|\vec{k}|\) of the lattice sites differs from the nominal shell
radius. Additionally, the measurement of \(L(t)\) in simulations with
periodic boundary conditions is subject to finite size effects. More
importantly, in suspensions of anisotropic particles, we cannot expect
that the structure factor is isotropic. Therefore, we also used the 3D
structure factor to calculate a lateral domain size in each Cartesian
direction from the second moments
\cite{jansen_bijels_2011,gunther_timescales_2014}
%
\begin{equation}
L_\beta(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}k_\beta^2 S(\vec{k},t)}} .
\end{equation}
%
This allows us to compute a domain size parallel
(\(L_\parallel=L_z\)) and perpendicular (\(L_\perp=(L_x+L_y)/2\)) to the
direction of \(\vec{B}\) separately. The average domain size \(L_d(t)\)
is computed as \(L_d(t)=\sum_\beta L_\beta(t) / 3\).
%
\begin{equation}
L_d(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}\sum_\beta k_\beta^2 S(\vec{k},t)}} .
\end{equation}
%
%In case the structure factor is isotropic
%\(S(\vec{k},t)=S(k,t)\), this is equivalent to the ratio of the fourth
%and second moments of the spherically averaged structure factor.
%
%\begin{equation}
%\begin{split}
%  L_d(t) &%= \left[\sum_{\beta}\left(\frac{1}{L_\beta(t)}\right)^2\right]^{-\frac{1}{2}}
%  = 2\pi \sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}\sum_\beta k_\beta^2 S(\vec{k},t)}} \\
%L_\perp(t) &%= \left[\left(\frac{1}{L_x(t)}\right)^2+\left(\frac{1}{L_y(t)}\right)^2\right]^{-\frac{1}{2}}
%= 2\pi \sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}} \left(k_x^2+k_y^2\right) S(\vec{k},t)}} \\
%L_\parallel(t) &= 2\pi \sqrt{\frac{\sum_{\vec{k}} S(\vec{k},t)}{\sum_{\vec{k}} k_z^2 S(\vec{k},t)}}
%\end{split}
%\end{equation}

Before calculating the structure factor \(S(\vec{k},t)\), we
first filled the lattice sites covered by particles with the average
density at the surrounding lattice sites. We employed an iterative
procedure that fills the particles layer by layer, starting at the
surface layer and using equation \eqref{eq:fill_particles} to
approximate the density at solid nodes that have a least one neighboring
fluid node. This is repeated until all solid nodes have been filled with
fluid. The order parameter \(\phi(\vec{x},t)\) was then calculated for
the filled density fields \(\rho_k(\vec{x},t)\). Filling the lattice
sites covered by particles is crucial to avoid spurious oscillations of
the structure factor at length scales corresponding to the particle
size.

@@

The anisotropy of domain size suggests that the magnetic field
influences the bijel morphology at the microscale. To investigate
whether these structural changes have an effect on macroscopic
transport properties, we also characterized the tortuosity of the
bijels. Various definitions of tortuosity have been proposed in the
literature \cite{dasilva_tortuosity_2022}, including geometric,
diffusional, and hydraulic tortuosity. Here we use the diffusional
tortuosity $\tau= \frac{\epsilon D}{D_{\text{eff}}}$, i.e., the ratio of the 
effective diffusivity $D_{\text{eff}}$ and the intrinsic diffusivity
$D$. Porous structures were generated from the order parameter field
at the last time step of the simulations by binarizing the raw order
parameter data using a threshold of zero. We masked the largest
connected component representing the percolating domain and then used
the Python package \texttt{taufactor}\cite{cooper_taufactor_2016} to
calculate the tortuosity in the three Cartesian coordinate
directions. \texttt{taufactor} determines the tortuosity by comparison
of steady-state diffusive flow through a porous medium with the bulk
diffusive flow in a control volume of the same size, molecular
diffusivity, and driving force \cite{cooper_taufactor_2016}. We used
the \texttt{PeriodicSolver} to employ periodic boundary conditions.

@@

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/results/paper1/particle_packing_viz.png}
\caption{Snapshots illustrating the particle packing at the interface for different particle shapes $\alpha$ with (left) and without (right) applied magnetic field $\bar{B}$. The right column shows the alignment of the symmetry axis of the oblate (top row) and prolate (bottom row) particles in the direction of the magnetic field indicated by arrows.}
\label{fig:packing_viz}
\end{figure}

@@

To measure the particle alignment quantitatively, we computed the
nematic order tensor \cite{veerman_phase_1992}
%
\begin{equation}
\tens{Q} = \frac{1}{n_p} \sum_i \left( \frac{3}{2} \hat{\vec{o}}_i \otimes \hat{\vec{o}}_i -\frac{1}{2}\mathsf{1} \right)
\end{equation}
%
where \(\hat{\vec{o}}_i\) is the orientation of the
\(i\)-th particles and \(n_p\) is the number of particles in the system.
The largest eigenvalue of $\tens{Q}$ yields the nematic order
parameter
%
\begin{equation}
S = \left\langle \frac{3}{2}\cos^2 (\theta) - \frac{1}{2} \right\rangle ,
\end{equation}
%
where
\(\theta=\arccos(\hat{\vec{o}}_i\cdot\hat{\vec{n}})\) is the angle
between the particle orientation and the nematic director
\(\hat{\vec{n}}\).

Figure \ref{fig:nematic_time} shows the time evolution of the nematic
order parameter \(S\) for the three different particle aspect ratios
\(\alpha\) and varying magnetic field strength \(\bar{B}\). The onset of
orientational order is nearly instantaneous and increases with
increasing field strength. For magnetic flux densities
\(\bar{B}\ge0.5\), the nematic order parameter saturates at
\(S\approx 1.0\) towards the end of the simulation. For the prolate
particles (\(\alpha=2.0\)), we observe a shoulder at around
\(10^4\Delta t\) and an intermittent plateau. The occurrence of this
plateau coincides with the slowed decay of the coarsening speed in Fig.
\ref{fig:coarsening_velocity}. This change in the orientational ordering
is even more distinct for the oblate particles, where the nematic order
decays intermittently before it increases again. The intermittent decay
for oblate particles, and the plateau for prolate particles, correlate
with the delayed onset of jamming in the direction perpendicular and
parallel to the magnetic field, respectively. These observations can be
interpreted as follows: The magnetic field aligns the particles with the
direction of the magnetic field early during the simulation. Due to this
alignment, the steric constraints between particles adsorbed at the
interface are reduced which allows further coarsening of the interfacial
area. However, during coarsening, the interfaces may re-orient
themselves and cause a capillary torque on the particles that competes
with the magnetic torque. This competition intermittently perturbs the
nematic order of the particles leading to the observed decay and plateau
of the nematic order parameter. Accordingly, the decay of \(S\) is most
pronounced at the weakest magnetic field \(\bar{B}=0.2\).

@@

% \section*{Data availability}

% The data and code used to generate the figures in this manuscript are available as an interactive notebook on Code Ocean at https://doi.org/00.0000/00000000. Due to size limitations, the raw simulation data are available for download on request from the authors.

% \section*{Acknowledgements}
% %The Acknowledgements come at the end of an article after Conflicts of interest and before the Notes and references.
% This work was supported by the National Science Foundation under NSF
% Awards DMR-1944942 and OIA-2346036. Any opinions, findings and
% conclusions or recommendations expressed in this material are those of
% the authors and do not necessarily reflect those of the National Science
% Foundation. This work used NCSA Delta at the National Center for
% Supercomputing Applications through allocation PHY220131 from the
% Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support
% (ACCESS) program, which is supported by National Science Foundation
% grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296. We
% wish to thank Jens Harting and Alexander Wagner for insightful
% discussions.

%%%END OF MAIN TEXT%%%

%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.

%\balance

%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
%\renewcommand\refname{Notes and references}

%%%REFERENCES%%%
% \bibliographystyle{rsc} % RSC's .bst file
% \bibliography{bijels}

>>>>

The addition of microstructure control in bijels can be utilized to improve the performance of porous materials
or soft matter derived from their co-continuous, tortuous microstructure. The importance of microstructure control
allows an increase in the charge/discharge rate of batteries through reducing the material tortuosity, or improving the 
adhesion and growth rate of bone grafts. These properties are imparted during the fabrication step of the bijel. Anisotropic
particles adsorbed at interfaces induce shape-dependent capillary interactions that have distinct effects on the
microstructure of emulsions, resulting in migration or rotation of particles at interfaces, in addition to the addition of
distinct timescales of domain coarsening. \cite{loudet_capillary_2005, cavallaro_curvature-driven_2011, gunther_lattice_2013}

In bijels, ellipsoidal stabilizers have been demonstrates to result in additional timescales of domain coarsening 
at long timescales owing to particle re-orientation at the interface due to interparticle capillary interactions 
at long timescales. \cite{gunther_timescales_2014} In addition to possessing more timescales, property, bijels stabilized 
with rod-like particles have been shown to result in smaller characteristic length scales, owing to the particles having a
larger cross sectional area resulting in jamming occurring at larger interfacial areas. \cite{hijnen_bijels_2015}

Magnetic response has been used in particle stabilized emulsions in the past, causing controlled emulsion failure and
on-demand droplet coalescence for encapsulation and cargo release applications. \cite{melle_pickering_2005,zhou_magnetic_2011}
Fields also enable self assembly through the control of interface deformation derived capillary forces.
\cite{morgan_understanding_2012,davies_interface_2014,davies_dipolar_2015} Magnetically responsive spherical particle 
stabilized bijels have thus far not yielded significant microstructural modifications. \cite{kim_bijels_2010} However, 
electric fields have been shown to cause percolating cylindrical domains in electrically responsive bijels. \cite{carmack_tuning_2018}
The electric field causes polar interactions between particles, which causes the formation of chains of particles that
also function as nucleation sites for phase separation. \cite{carmack_tuning_2018}

In this chapter, we investigate the effect of external magnetic fields on the formation of bijels stabilized by 
anisotropic magnetic particles. The microstructure of bijels stabilized with oblate, spherical and prolate particles, 
namely the characteristic length scale and tortuosity, will be assessed as a function of the applied magnetic field
strength. We also characterize the underlying mechanisms governing the observed behavior.

% Anisotropic particles, when adsorbed at liquid interfaces, induce
% shape-dependent capillary interactions that can have a distinct effect
% on the microstructure of emulsions
% \cite{loudet_capillary_2005,loudet_self-assembled_2009}. For instance,
% Madivala et
% al.~\cite{madivala_exploiting_2009,madivala_self-assembly_2009} found
% that emulsion stability can be significantly enhanced by ellipsoidal
% particles. Additionally, capillary interactions can lead to direct
% assembly of anisotropic particles
% \cite{bowden_self-assembly_1997,lewandowski_orientation_2010,stebe_oriented_2009,botto_capillary_2012}.
% Ellipsoidal particles can rotate and migrate on curved surfaces
% \cite{cavallaro_curvature-driven_2011}, where the interface curvature
% acts as an external field that controls the orientation and placement
% of particles \cite{furst_directing_2011}. Anisotropic particles at
% fluid interfaces were considered in Monte Carlo simulations by Bresme
% et al. to test the accuracy of thermodynamic models for the free
% energy as a function of the particle-fluid interactions, particle
% size, and orientation
% \cite{bresme_orientational_2007,bresme_computer_2008}. Harting and
% co-workers \cite{gunther_lattice_2013,gunther_timescales_2014} have
% performed lattice Boltzmann simulations of anisotropic particles at
% liquid interfaces and in emulsions. They found that anisotropic
% particles affect the dynamics of emulsion formation by introducing two
% additional timescales associated with particle rotation. Following
% initial adsorption, particles rotate to align their larger
% cross-section with the interface. When jamming sets in, capillary
% interactions cause the particle to re-orient in the interface leading
% to prolonged coarsening at long timescales. Due to the larger aspect
% ratio, ellipsoidal particles can stabilize a larger interface area
% leading to smaller domain sizes. Hijnen et
% al.~\cite{hijnen_bijels_2015} experimentally realized bijels
% stabilized by colloidal rods and analyzed their effect on the
% interface and domain morphology. Their experiments confirmed that
% jamming occurs at lower interface separations than for spherical
% particles with the same volume fraction. Moreover, they observed that
% the overall structure remains isotropic and attributed this to the
% rapid jamming of the rods. Additionally, they found that some rods
% tilted out of the interface due to the compression induced by domain
% coarsening.

% Particle-stabilized emulsions can be made responsive to external
% magnetic fields by magnetic particles
% \cite{melle_pickering_2005,zhou_magnetic_2011}. An external field can
% then be used to control the stability of the emulsion
% \cite{melle_pickering_2005} and the self-assembly of particles
% \cite{dassanayake_structure_2000,leunissen_directing_2009}. The tendency
% of anisotropic particles to adopt distinct orientations gives rise to an
% orientation transition when an external field is applied
% \cite{morgan_understanding_2012,davies_interface_2014,davies_dipolar_2015}.
% Magnetic fields thus offer a means to control the capillary interactions
% of ellipsoidal particles at an interface. For bijels stabilized by
% spherical magnetic particles, simulations by Kim et
% al.~\cite{kim_bijels_2010} revealed that magnetic fields do not have a
% significant effect on the microstructure of bijels. In contrast, Carmack
% and Millett \cite{carmack_tuning_2018} demonstrated that the structure
% of thin-film bijels stabilized by dielectric particles can change
% drastically when an external electric field is applied. The dielectric
% contrast of the particles induces polar interactions that cause liquid
% domain alignment leading to percolating cylindrical domains in the
% thin-film bijel. In addition, they found that particle chains can act as
% nucleation sites for phase separation. This suggests that anisotropic
% interactions can have a considerable effect on the phase separation and
% the resulting emulsion morphology and may offer a means to control the
% size and shape of liquid domains.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/results/paper1/microstructure_viz.png}
    \caption{Snapshots of emulsion gels stabilized by prolate ellipsoids after $t=5\cdot10^3$ timesteps (left) and after $10^5$ timesteps (right). The top row shows the structure forming without a magnetic field. 
             The bottom row shows the structure forming in an applied magnetic field of reduced strength $\bar{B}=1.0$. The snapshots are colored according to the order parameter $\phi$.}
    \label{fig:microstructure_viz}
\end{figure}

@@

In LB simulations of spinodal decomposition, a characteristic length
scale can be obtained from the structure factor, defined in Section \ref{section:domain_size}. \cite{kendon_3d_1999,kendon_inertial_2001} 
It is worth noting that this definition of average domain size does not necessarily yield values that coincide with the typical
domain size observed visually in snapshots such as those shown in Fig.~\ref{fig:microstructure_viz}. However, in a dynamic
scaling regime, different measures of domain size are expected to differ only by a constant ratio. We found that the commonly 
used measure \(L_1(t)\) tends to overestimate the observed domain size. Numerically, the calculation of the spherically 
averaged structure factor is subject to poor statistics in the low \(k\)-shells, where the average \(|\vec{k}|\) of the 
lattice sites differs from the nominal shell radius. Additionally, the measurement of \(L(t)\) in simulations with
periodic boundary conditions is subject to finite size effects. More importantly, in suspensions of anisotropic particles, 
we cannot expect that the structure factor is isotropic. Therefore, we also used the 3D structure factor to calculate a lateral domain 
size in each Cartesian direction from the second moments also defined in Section \ref{section:domain_size}. 
\cite{jansen_bijels_2011,gunther_timescales_2014} We add the additional processing calculating a parallel and perpendicular domain size
from the results in each cartesian direction, defining $L_{\parallel}=L_z$ and $L_{\perp} = \frac{L_x+L_y}{2}$

Before calculating the structure factor \(S(\vec{k},t)\), we first filled the lattice sites covered by particles with the average
density at the surrounding lattice sites as detailed in Section \ref{section:filling_routine}. We plot the structure factor of the structures in
Figure \ref{fig:structure_factor}.

@@

The anisotropy of domain size suggests that the magnetic field
influences the bijel morphology at the microscale. To investigate
whether these structural changes have an effect on macroscopic
transport properties, we also characterized the tortuosity of the
bijels. Various definitions of tortuosity have been proposed in the
literature \cite{dasilva_tortuosity_2022}, including geometric,
diffusional, and hydraulic tortuosity. Here we use the diffusional
tortuosity $\tau= \frac{\epsilon D}{D_{\text{eff}}}$, i.e., the ratio of the 
effective diffusivity $D_{\text{eff}}$ and the intrinsic diffusivity
$D$. We calculate the tortuosity using the procedure detailed in Section \ref{section:tortuosity}

@@

\subsection{Curvature}

Beyond the length scale, curvature analysis offers insights into the
local topological landscape of the bijel, facilitating further
microstructure optimization for various applications.
\cite{reeves_quantitative_2016} We plot the area averaged mean and
gaussian curvature \(H\Sigma^{-1}\) and \(K\Sigma^{-2}\) for bijels
stabilized with oblate, spherical and prolate ellipsoids in Figure
\ref{fig:curvature-vs-B_ss}.

\begin{figure} 
    \centering 
    \includegraphics[width=\textwidth]{figures/results/paper1/curvature-vs-B_ss.png} 
    \caption{Plot of the area averaged mean $\langle H \rangle \Sigma^{-1}$ and Gaussian 
            $\langle K \rangle \Sigma^{-2}$ curvature in the left and right respectively. Each 
            plot contains the $\langle H \rangle \Sigma^{-1}$ and $\langle K \rangle \Sigma^{-1}$ 
            at the final timestep averaged across 3 runs for particles with $\alpha = 0.5, 1, 2$. 
            We observe that the mean curvature does not demonstrate any trends, while the Gaussian 
            curvature shows a reduction as the magnetic field strength is increased for 
            ellipsoidal particles.} 
    \label{fig:curvature-vs-B_ss}
\end{figure}

Error bars are represented by the standard
deviation of the three simulations per parameter set performed. The area
averaged mean curvature shows little dependence on the applied field
strength with values around 0 and within error of one another. A value
near or at 0 is expected as the fluid volume fraction is equal and the
particles are neutrally wetting, implying the interface should not have
a preferred direction of curvature. \cite{jinnai_interfacial_2001} For
ellipsoidal particles, the gaussian curvature becomes more negative as
larger field strengths are applied. We also see that the gaussian
curvature becomes more positive as the interfacial area of the particle
increases. The trend of these results are inverse to the interfacial
area of each particle, \(A_{rm}\). This is expected behavior and is in
agreement with experiments that have investigated the effect of particle
size on the curvature of bijels. \cite{reeves_quantitative_2016}

Reeves and coworkers identified that nanoparticles were better able to
lock in the spinodal structure of the liquid-liquid phase separation as
they were less disruptive the the curvature of the interface due to
their lower interfacial adsorption energy. This effect can be observed
as a more negative Gaussian curvature as a result of the greater number
of hyperbolic surfaces, a hallmark of a minimal surface. In our system,
the oblate and prolate particles have 60\% and 30\% more interfacial
area than the spherical particles, resulting in the ellipsoidal
particles being more disruptive to the curvature of the interface,
observed as a more positive Gaussian curvature. As the magnetic field
strength is increased, the packing of the particles at the interface
changes, as observed from the radial distribution function.
\cite{karthikeyan_formation_2024} These changes in particle packing
change the way the particles affect the curvature of the interface,
resulting in the curvature changes observed. To investigate the field
dependence on the curvature, we plot the area averaged Gaussian
curvature as a function of time.

\begin{figure} 
    \centering 
    \includegraphics[width=\textwidth]{figures/results/paper1/curvature_vs_coverage.png} 
    \caption{Plots of Gaussian $\langle K \rangle \Sigma^{-2}$ curvature over time averaged across 3 runs. 
    From left to right, we have particles with $\alpha = 0.5, 1, 2$. We observe that the initial stages of 
    evolution are unaffected by the applied magnetic field. However, after about 10000 timsteps, we see a 
    magnetic field dependence on the evolution of the Gaussian curvature. The preservation of this divergence 
    occurs when ellipsoidal particles are used but not spherical particles.} 
    \label{fig:curvature_vs_coverage}
\end{figure}

Given we did not see any noteworthy changes in the mean curvature in
Figure \ref{fig:curvature-vs-B_ss}, we focused on the Gaussian curvature
in Figure \ref{fig:curvature_vs_coverage}. We observe that for all
simulated systems, the Gaussian curvature increases over time until an
invariance is reached near the end of the simulation, signifying the
bijel microstructure is locked in place. We also observe that for all
microstructures the curvature over time does not increase at the same
rate, characterized with dips at 10 to 20,000 timesteps for all
particles and a magnetic field dependent change in the curvature
increase.

However, for the spherical particles this local topology change
disappears once the spinodal decomposition ceases, observed as a
convergence of all curvatures. This is not observed for the bijels
stabilized by ellipsoidal particles. Despite all our curvature
calculations being normalized with their surface area to volume ratio,
we still see curvature differences as a function of the applied magnetic
field. Experiments have identified that curvature changes can occur due
to the formation of monogels, which are particle networks formed from
bijels that can withstand the remixing of their constituent liquids.
\cite{sanz_colloidal_2009, lee_making_2013} This process cannot occur in
our simulations as we are utilizing repulsive particles. Therefore, the
change in curvature arises from the differences in local state of the
colloids at the interface.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/results/paper1/particle_packing_viz.png}
\caption{Snapshots illustrating the particle packing at the interface for different particle shapes $\alpha$ with (left) 
        and without (right) applied magnetic field $\bar{B}$. The right column shows the alignment of the symmetry axis of 
        the oblate (top row) and prolate (bottom row) particles in the direction of the magnetic field indicated by arrows.}
\label{fig:packing_viz}
\end{figure}

@@

To measure the particle alignment quantitatively, we computed the
nematic order tensor using the method detailed in Section \ref{section:nematic_order_parameter} 
Figure \ref{fig:nematic_time} shows the time evolution of the nematic
order parameter \(S\) for the three different particle aspect ratios
\(\alpha\) and varying magnetic field strength \(\bar{B}\). The onset of
orientational order is nearly instantaneous and increases with
increasing field strength. For magnetic flux densities
\(\bar{B}\ge0.5\), the nematic order parameter saturates at
\(S\approx 1.0\) towards the end of the simulation. For the prolate
particles (\(\alpha=2.0\)), we observe a shoulder at around
\(10^4\Delta t\) and an intermittent plateau. The occurrence of this
plateau coincides with the slowed decay of the coarsening speed in Fig.
\ref{fig:coarsening_velocity}. This change in the orientational ordering
is even more distinct for the oblate particles, where the nematic order
decays intermittently before it increases again. The intermittent decay
for oblate particles, and the plateau for prolate particles, correlate
with the delayed onset of jamming in the direction perpendicular and
parallel to the magnetic field, respectively. These observations can be
interpreted as follows: The magnetic field aligns the particles with the
direction of the magnetic field early during the simulation. Due to this
alignment, the steric constraints between particles adsorbed at the
interface are reduced which allows further coarsening of the interfacial
area. However, during coarsening, the interfaces may re-orient
themselves and cause a capillary torque on the particles that competes
with the magnetic torque. This competition intermittently perturbs the
nematic order of the particles leading to the observed decay and plateau
of the nematic order parameter. Accordingly, the decay of \(S\) is most
pronounced at the weakest magnetic field \(\bar{B}=0.2\).

@@
